Project 2

### 2015-03-31 project2/paper/main.tex
Adding a reference to the application
Finalizing the current draft

\section{Previous Work}
Many large cities have rapid transit systems, so considerable research time has been invested on building reliable
subways. Commercial software such as OpenTrack \cite{opentrack} have been developed, and even video games such as Train
Simulator \cite{trainsimulator} and Open Rails \cite{openrails}. Algorithms for handling single, double and tracks exist, such as the basic intuitive
algorithm for sharing tracks between trains presented by Fiorini and Botter \cite{fioroni}. Dessouky and Leachman
\cite{dessouky_leachman_95} provide an event-driven algorithm for single and double track rail networks, incorporating
acceleration. Lu et al \cite{quan_lu} expand the model to triple tracks while enhancing the acceleration model to
handle multiple track speed limits.

@@

Because of this, we can represent the entire metro as queuing network with sections between stations as the queue. In
our case, we will only be modeling the section of track between Rosslyn and Stadium-Armory where the Orange, Blue, and
Silver share a track (See \ref{fig:metromap}). Since there are no track crossovers, we can consider these three lines
to be isolated from other lines.

@@

The API provides data by station with an estimate at each station of the wait time until the next train. For instance,
the API for a station on the blue line might list a train to Franconia-Springfield (westbound) in 1 minute. If there
also happens to be an Orange that passes through that station, you may also see an estimate for the next Orange (if it
close enough).

Given this information, we can calculate how often a train runs simply by marking the timestamp of the API call and
offset by the estimated time until arrival provided by the API. For instance, if we query a station at 12:00 PM, and
the API estimates a train arriving in 3 minutes. We can roughly estimate that the train will arrive at 12:03. This
gives us our arrival rate Furthermore, we can query by train line as well to verify that trains run according to the
schedule that WMATA provides on its website \cite{metroschedule}.

@@

We can still calculate an estimated travel time through the API as well. Given two neighboring stations $A$ and $B$, if
we query $A$ for an upcoming trains of color $c$ in the direction of $B$, we will get some estimate for the arrival. We
can query $B$ for the same color train in the same direction to obtain a second estimate. Subtracting these two numbers
provides an estimate of the travel time between stations. This is making an assumption that there is no train of the
same color in the middle that may lower the estimate. However, if we restrict our queries on $A$ to trains that are only
a minute away, we are unlikely to have that problem since the stations are so close.

@@

Train interarrivals were measured by line and direction. \ref{fig:eastboundblueinterarrivals} demonstrates a case where
the data appears to be normal. However, other cases such as \ref{fig:westboundsilverinterarrivals} show that this is
not necessarily true. Graphs for the remaining directions and line colors were omitted for brevity, but are included in
the submission. Due to time constraints, we'll use an empirical distribution rather than trying to fit anything.

@@

\section{Caveats and Limitations}
With any sort of simulation there are simplications and assumptions that are made, and this project is of course no
different. In some cases the reduction in detail does not harm accuracy but simply makes the model designer's and
programmer's life easier. However, sometimes this is not the case and the simplifications greatly reduce the accuracy
of the model. Thus, it is the modeler's job to judiciously choose which details to omit and leave in when building the
model.

One of the simplifications that we've made is that we've excluded all other stations in the metro system except those
between Rosslyn and Stadium Armory. Other lines, the Red, Green, and Yellow lines, do not share track with any train,
nor do they cross at any point. In this respect, they are isolated (regardless of the fact that they may share
stations). However, ignoring Blue, Orange, and Silver stations beyond Stadium-Armory and Rosslyn could potentially
reduce model accuracy here because train events at those stations can very easily influence the stations within D.C.
We have chosen instead to model those exterior stations as a probability distribution which is based off of empirical
data, but we still run a risk.

@@

\section{Current and Remaining Work}
Currently, the app is simply mapping stations and labeling them. The entirety of data analysis is complete, and can be
viewed within the \texttt{queries} directory, along with the \texttt{utils}. A few data files are strewn about the
codebase as well. The WMATA API query code is stored within the \texttt{wmata} folder.

I have also copied some code over from the first project since it will be applicable, such as a discrete event queue.
However, the \texttt{Track} class remains to be written and is integral to the application. Furthermore, the actual
train scheduling algorithm itself needs to be written.

>>>>

\section{Previous Work}
Many large cities have rapid transit systems, so considerable research time has been invested on building reliable
subways. Commercial software such as OpenTrack \cite{opentrack} have been developed, and even video games such as Train
Simulator \cite{trainsimulator} and Open Rails \cite{openrails}. Algorithms for handling single, double and triple tracks exist,
such as a basic
algorithm for sharing tracks between trains presented by Fiorini and Botter \cite{fioroni}. Dessouky and Leachman
\cite{dessouky_leachman_95} provide an event-driven algorithm for single and double track rail networks, incorporating
acceleration. Lu et al \cite{quan_lu} expand the model to triple tracks while enhancing the acceleration model to
handle multiple track speed limits.

@@

Because of this, we can represent the entire metro as queuing network with sections between stations as the queue. In
our case, we will only be modeling the section of track between Rosslyn and Stadium-Armory where the Orange, Blue, and
Silver share a track (See \ref{fig:metromap}). Since there are no track crossovers, we can consider these three lines
to be isolated from other lines. However, we do need to keep in mind that for two tracks, each track is devoted to a
specific direction. Only in the case of a single track is when a track will handle both directions.

Because this will be a discrete-event simulation, we will need to be mindful of the events that travel through our
application. A first example that comes to mind is the \texttt{train:arrived} event, which will be received by the
Rosslyn and Stadium-Armory stations when a train arrives. This event will need to contain which line, along with the
direction its heading (although that is implied by the station in our case).

Track sections will need to inform neighboring stations when they are free, or if they are occupied. This of course
will influence when and where trains are sent in order to reach their destination. In this respect, track sections are
the queues in this network.

@@

The API provides data by station with an estimate at each station of the wait time until the next train. For instance,
the API for a station on the blue line might list a train to Franconia-Springfield (westbound) in 1 minute. If there
also happens to be an Orange that passes through that station, you may also see an estimate for the next Orange (if it
is close enough).

Given this information, we can calculate how often a train runs simply by marking the timestamp of the API call and
offset by the estimated time until arrival provided by the API. For instance, if we query a station at 12:00 PM, and
the API estimates a train arriving in 3 minutes, we can roughly estimate that the train will arrive at 12:03. This
gives us our arrival rate. Furthermore, we can query by train line as well to verify that trains run according to the
schedule that WMATA provides on its website \cite{metroschedule}.

@@

We can still calculate an estimated travel time between two stations through the API as well. Given two neighboring stations $A$ and $B$, if
we query $A$ for an upcoming trains of color $c$ in the direction of $B$, we will get some estimate for the arrival. We
can query $B$ for the same color train in the same direction to obtain a estimate of the arrival time at $B$.
Subtracting these two numbers provides an estimate of the travel time between stations. This makes an assumption that
there is no train of the same color in the middle that may skew the estimate. However, we can minimize the chance of
this happening if we restrict our queries on $A$ to trains that are only a minute away. Given that the stations are so
close, we minimize the likelihood that there is a ``middle train'' that causes inaccurate estimates.

@@

Train interarrivals were measured by line and direction. \ref{fig:eastboundblueinterarrivals} demonstrates a case where
the data appears to be normal. However, other cases such as \ref{fig:westboundsilverinterarrivals} show that this is
not necessarily true. Graphs for the remaining directions and line colors were omitted for brevity, but are included in
the submission. Due to time constraints, we'll use an empirical distribution rather than trying to fit any particular
distribution.

@@

\section{Caveats and Limitations}
With any sort of simulation of a real-world system, there are simplications and assumptions that are made, and this project is of course no
different. In some cases the reduction in detail does not harm accuracy but simply makes the model designer's and
programmer's life easier. However, sometimes this is not the case and the simplifications greatly reduce the accuracy
of the model. Thus, it is the modeler's job to judiciously choose which details to omit and leave in when building the
model.

One of the simplifications that we've made is that we've excluded all other stations in the metro system except those
between Rosslyn and Stadium Armory. Other lines, the Red, Green, and Yellow lines, do not share track with any train,
nor do they cross at any point. In this respect, they are isolated (regardless of the fact that they may share
stations). So, in this case, the simplification hinders accuracy in no way whatsoever. However, ignoring Blue, Orange,
and Silver stations beyond Stadium-Armory and Rosslyn could potentially
reduce model accuracy here because train events at those stations can very easily influence the stations within D.C.
We have chosen instead to model those exterior stations as a probability distribution which is based off of empirical
data, but we still run a risk that we may have not captured an important interaction.

@@

\section{Current and Remaining Work}
Currently, the app is simply mapping stations and labeling them based on information provided by the API. The entirety
of data analysis is complete, and can be
viewed within the \texttt{queries} directory, along with the \texttt{utils}. A few data files are strewn about the
codebase as well. The WMATA API query code is stored within the \texttt{wmata} folder.

I have also copied some code over from the first project since it will be applicable, such as a discrete event queue.
However, the \texttt{Track} class remains to be written and is integral to the application, along with the \texttt{Train}
class. Furthermore, the actual
train scheduling algorithm itself needs to be written. However, the current application is deployed here
\cite{aprilandchip}.
